### 2019-10-22 Writing/Palaeontology/draft.tex
Updated figures numbers for SOM3

\subsection{Comparing topologies}
We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 7 and 8).
Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
First, in biology, this tree is always unknown.
Second, our objective is to measure the direct effect of character correlation approximated by the difference in topology between the ``normal'', ``maximised'' and ``minimised'' trees.
When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 12, 13 and 14.

@@

\subsection{Effect of character differences on topology}
As expected, there is a significant effect of character difference in the ability to recover the correct topology (in terms of clades).
The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
Intuitively, one would hypothesise that the ``minimised'' character difference scenario would lead to incorrect but consistent trees (i.e. because the characters are not different, they all support the same topology that might be incorrect), the ``maximised'' scenario to lead to poorly resolved trees (i.e. because characters are different, the matrix has a lot of homoplasy leading to badly resolved topologies) and the ``randomised'' scenario to perform the best at recovering the correct topology (i.e. a simple bootstrap procedure). 
However, throughout our simulation, the ``minimised'' character difference scenario (i.e. increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario (i.e. decreased correlation) led to slightly worst trees (Fig. \ref{Fig:RF_results_best}).
This effect of increasing or reducing character difference did not however appear to affect the taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
However this results needs to be contrasted with the fact that the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 9, 10 and 11).
Furthermore, it is important to note that on average, fewer characters were duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 4, 5 and 6 - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 1, 2 and 3).
This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.

@@

The differences in performance of the Bayesian and parsimony approaches observed here could be due to the inherent mechanisms of each method.
For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate scores differently:
(1) in parsimony, the topology will probably be given a bad score (one that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the focal topology.
The tree search will thus likely result in a topology island that will not contain the given topology $T$.
(2) in Bayesian inference, the topology will also be given a bad score (i.e. low likelihood) although the high homoplasy can be accommodated in the tree through high evolution rates or/and long branches.
The rate and the branch length being two parameters among others, the score (the likelihood) will change less drastically than for using parsimony.
Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
This difference in both mechanisms could explain why, on average, Bayesian inference seems better to recover the ``normal'' topology than maximum parsimony.
This can be also supported by the fact that Bayesian inference recovered the ``true'' tree topology slightly but consistently better (supplementary materials 3, figures 12, 13, and 14).

@@

Second, measuring and modifying character correlation is difficult. 
In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation (supplementary materials 3, figures 1, 2 and 3).
In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.
This could lead to correlations between characters to be more present in some parts of the trees that other \citep[e.g. in the case of inapplicable data ][]{BrazeauNA}.
However, because of the number of characters, it is actually complex to actually measure their correlation in a biological sense and is still actively discussed in the literature \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,goswami2006morphological,goswami2010,goswami2014macroevolutionary,Grabowski2016}.
Additionally, as discussed in the introduction, character correlation can also simply arise by chance due to the discrete coding scheme (i.e. some sets of characters can be highly correlated but effectively describe independent information).
Therefore, we made the choice to simplify our simulations by generating character correlation as a stochastic process rather than a biological one.

>>>>

\subsection{Comparing topologies}
We compared the topologies using the same approach as in \cite{Guillerme2016146}: we measured both the Robinson-Foulds distance \citep{RF1981} and the triplets distance \citep{dobson1975triplets} between the trees inferred from the ``maximised'', ``minimised'' and ``randomised'' matrices and the tree inferred from the ``normal'' matrix.
We explored the effect of character difference on recovering the ``normal'' topology by comparing the ``maximised'', ``minimised'' and ``randomised'' trees to the ``normal'' tree (Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best} and supplementary materials 3 Figs 10 and 11).
Note that we are not comparing the trees to the ``true'' tree used to simulate the matrices.
First, in biology, this tree is always unknown.
Second, our objective is to measure the direct effect of character correlation approximated by the difference in topology between the ``normal'', ``maximised'' and ``minimised'' trees.
When measuring the difference between these trees and the ``true'' tree, we would also confound the effect of simulating a birth-death tree and simulating a discrete morphological matrices from it.
However, we did report the comparison between the ``true'' trees and both the ``normal'' and ``randomised'' in the supplementary materials 3, Figs 15, 16 and 17.

@@

\subsection{Effect of character differences on topology}
As expected, there is a significant effect of character difference in the ability to recover the correct topology (in terms of clades).
The character difference metric can be seen as the inverse of character correlation (see Methods above): a high character difference approximates a low level of character correlation and vice versa.
When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
Intuitively, one would hypothesise that the ``minimised'' character difference scenario would lead to incorrect but consistent trees (i.e. because the characters are not different, they all support the same topology that might be incorrect), the ``maximised'' scenario to lead to poorly resolved trees (i.e. because characters are different, the matrix has a lot of homoplasy leading to badly resolved topologies) and the ``randomised'' scenario to perform the best at recovering the correct topology (i.e. a simple bootstrap procedure). 
However, throughout our simulation, the ``minimised'' character difference scenario (i.e. increased correlation) tended to result in trees slightly closer to the ``true'' tree and the ``maximised'' character difference scenario (i.e. decreased correlation) led to slightly worst trees (Fig. \ref{Fig:RF_results_best}).
This effect of increasing or reducing character difference did not however appear to affect the taxon placement ($NTS_{Tr}$) for which the confidence intervals overlapped too much to be able to discern clear statistical differences (Fig.\ref{Fig:Tr_results_best}).
However this results needs to be contrasted with the fact that the trees generated by the ``minimised'' character difference scenario do not appear better resolved (towards any topology) than the other scenarios (see supplementary material 3, Figs 12, 13 and 14).
Furthermore, it is important to note that on average, fewer characters were duplicated (replaced) in the ``minimised'' matrices compared to both the ``maximised'' and ``randomised'' matrices (supplementary material 3, Figs 7, 8 and 9 - likely due to the fact that replacing only a few highly different characters can drastically decrease the character difference).
However, this change in the amount of replaced characters did not change the way average character differences varied across the three transformed matrices: the character differences were always consistently increased (``maximised''), decreased (``minimised'') and maintained at a normal level (``randomised'') for each replicate (supplementary material 3, Figs 4, 5 and 6).
This suggest that character difference can be decreased with a smaller amount of characters replacements than increasing it.

@@

The differences in performance of the Bayesian and parsimony approaches observed here could be due to the inherent mechanisms of each method.
For any given topology $T$ that was obtained from the ``normal'' matrix and a matrix with high homoplasy, both methods will generate scores differently:
(1) in parsimony, the topology will probably be given a bad score (one that implies many changes along the tree) and the optimality criterion (favouring the minimum score) will likely discard the focal topology.
The tree search will thus likely result in a topology island that will not contain the given topology $T$.
(2) in Bayesian inference, the topology will also be given a bad score (i.e. low likelihood) although the high homoplasy can be accommodated in the tree through high evolution rates or/and long branches.
The rate and the branch length being two parameters among others, the score (the likelihood) will change less drastically than for using parsimony.
Furthermore, in Bayesian inference, a reasonable difference in optimality between two topologies (the acceptance) will not necessarily mean that the given topology $T$ will be discarded.
This difference in both mechanisms could explain why, on average, Bayesian inference seems better to recover the ``normal'' topology than maximum parsimony.
This can be also supported by the fact that Bayesian inference recovered the ``true'' tree topology slightly but consistently better (supplementary materials 3, figures 15, 16, and 17).

@@

Second, measuring and modifying character correlation is difficult. 
In our simulation protocol we chose to create simulation by duplicating characters in a matrix to maximise or minimise correlation (supplementary materials 3, figures 4, 5 and 6).
In biology, this correlation arises from either intra-organismal or evolutionary mechanisms.
This could lead to correlations between characters to be more present in some parts of the trees that other \citep[e.g. in the case of inapplicable data ][]{BrazeauNA}.
However, because of the number of characters, it is actually complex to actually measure their correlation in a biological sense and is still actively discussed in the literature \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,goswami2006morphological,goswami2010,goswami2014macroevolutionary,Grabowski2016}.
Additionally, as discussed in the introduction, character correlation can also simply arise by chance due to the discrete coding scheme (i.e. some sets of characters can be highly correlated but effectively describe independent information).
Therefore, we made the choice to simplify our simulations by generating character correlation as a stochastic process rather than a biological one.
